---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
editor_options:
  chunk_output_type: inline
---

```{r}
pacman::p_load(tidyverse, 
               gtsummary, 
               janitor, 
               NHANES, 
               palmerpenguins)
```

# Recapitulation

## Data exploration

```{r}
head(penguins)
```
```{r}
summary(penguins)
```

### Explore NA

```{r}
pacman::p_load(visdat)
```

```{r}
penguins %>% 
  visdat::vis_dat()
```


### Data visualization

#### Points
```{r}
penguins %>% 
  ggplot(aes(x = bill_length_mm, 
             y = bill_depth_mm)) + 
  geom_point() 
```

##### Add layers

```{r}
penguins %>% 
  ggplot(aes(x = bill_length_mm, 
             y = bill_depth_mm)) + 
  geom_point() + 
  aes(color = sex) 
```

##### Facetting

```{r}
penguins %>% 
  ggplot(aes(x = bill_length_mm, 
             y = bill_depth_mm)) + 
  geom_point() + 
  aes(color = sex) + 
  facet_wrap(~ species)
```

### One variable continuous

```{r}
penguins %>%
  filter(species == "Adelie") %>%
  drop_na() %>% 
  ggplot(aes(x = body_mass_g)) +
  geom_histogram(bins = 12) +
  facet_grid(sex ~ .) + 
  labs(
    title = "Adelie body mass", 
    x = "g", 
    y = "Count"
  )
```
Change theme

```{r}
penguins %>%
  filter(species == "Adelie") %>%
  drop_na() %>% 
  ggplot(aes(x = body_mass_g)) +
  geom_histogram(bins = 12) +
  facet_grid(sex ~ .) + 
  labs(
    title = "Adelie body mass", 
    x = "g", 
    y = "Count"
  ) + 
  theme_minimal()
```
```{r}
pacman::p_load(ggthemes)
```

```{r}
penguins %>%
  filter(species == "Adelie") %>%
  drop_na() %>% 
  ggplot(aes(x = body_mass_g)) +
  geom_histogram(bins = 12) +
  facet_grid(sex ~ .) + 
  labs(
    title = "Adelie body mass", 
    x = "g", 
    y = "Count"
  ) + 
  theme_minimal() + 
  ggthemes::theme_economist()
```

### Two variables continuous

```{r}
penguins %>% 
  drop_na() %>% 
  ggplot(aes(x = body_mass_g, 
             y = bill_depth_mm)) + 
  geom_point() + 
  aes(color = sex) 
```
#### Explore correlations

```{r}
penguins %>% 
  drop_na() %>% 
  ggplot(aes(x = body_mass_g, 
             y = bill_depth_mm)) + 
  geom_point() + 
  aes(color = sex) +
  facet_grid(sex ~ species) + 
  geom_smooth()
```

### Two variables continuous vs nominal

```{r}
penguins %>% 
  drop_na() %>% 
  ggplot(aes(x = species, 
             y = body_mass_g, 
             fill = species)) + 
  geom_boxplot()
```

### Trends

```{r}
gapminder <- gapminder::gapminder
```

```{r}
head(gapminder)
```
First, revise DATA WRANGLING

## Data Wrangling

### Filter

```{r}
gapminder %>% 
  filter(country == "Germany")
```
```{r}
gapminder %>% 
  filter(continent == "Europe" & year == 2007)
```


### Select
```{r}
gapminder %>% 
  select(continent, year, lifeExp)
```

### Group_by

```{r}
gapminder %>% 
  group_by(continent, year)
```


### Summarize

```{r}
gapminder %>% 
  group_by(continent, year) %>% 
  summarise(LifeExpMean = mean(lifeExp))
```
```{r}
pen
```


### Pivoting

```{r}
unicef <- read_csv("https://bit.ly/unicef-wide")
```
```{r}
head(unicef)
```

Clean the names
```{r}
unicef <- unicef %>% 
  janitor::clean_names()
```





#### Longer

```{r}
unicef %>% 
  pivot_longer(u5mr_1950:u5mr_2015, 
               names_to = "year", 
               values_to = "values")
```
`
Save a new dataframe

```{r}
unicef_long <- unicef %>% 
  pivot_longer(u5mr_1950:u5mr_2015, 
               names_to = "year", 
               values_to = "values")
```

filter some countries



```{r}
unicef_long %>% 
  filter(country_name %in% c("Latvia", "Lithuania", "Estonia")) %>% 
  ggplot(aes(x = year, 
             y = values, 
             color = country_name)) + 
  geom_point()
```

```{r}
unicef_long %>% 
  filter(country_name %in% c("Latvia", "Lithuania", "Estonia")) %>% 
  ggplot(aes(x = year, 
             y = values, 
             color = country_name)) + 
  geom_line()
```

```{r}
unicef_long %>% 
  filter(country_name %in% c("Latvia", "Lithuania", "Estonia")) %>% 
  ggplot(aes(x = year, 
             y = values, 
             color = country_name, 
             group = country_name)) + 
  geom_line()
```

#### Wider

```{r}
gapminder %>% 
  group_by(continent, year) %>% 
  summarise(LifeExpMean = mean(lifeExp))
```
```{r}
gapminder %>% 
  group_by(continent, year) %>% 
  summarise(LifeExpMean = mean(lifeExp)) %>% 
  pivot_wider( names_from = year, 
               values_from = LifeExpMean)
```
```{r}
gapminder %>% 
  filter(year %in% c(1952, 1972, 1992, 2002)) %>% 
  group_by(continent, year) %>% 
  summarise(LifeExpMean = mean(lifeExp)) %>% 
  pivot_wider( names_from = year, 
               values_from = LifeExpMean)
```


### Join datasets


Data from 

Short
https://datahub.io/core/country-list

Full
https://datahub.io/JohnSnowLabs/country-and-continent-codes-list#resource-country-and-continent-codes-list-csv


```{r}
countries_short <- read_csv("https://pkgstore.datahub.io/core/country-list/data_csv/data/d7c9d7cfb42cb69f4422dec222dbbaa8/data_csv.csv")


countries_list <- read_csv("https://datahub.io/JohnSnowLabs/country-and-continent-codes-list/r/country-and-continent-codes-list-csv.csv")
```
```{r}
head(countries_short)
```

Lets merge
First try
```{r}
left_join(unicef, countries_short, by = c("country_name" = "Name")) 
```


if ok, proceed
```{r}
unicef <- left_join(unicef, countries_short, by = c("country_name" = "Name")) 
```

```{r}
head(unicef)
```

```{r}
unicef %>% 
  relocate(Code, before = country_name)
```
```{r}
unicef <- unicef %>% 
  relocate(Code, before = country_name)
```


```{r}
head(unicef)
```


now add the continent

```{r}
head(countries_list)
```


```{r}
left_join(unicef, countries_list, by = c("Code" = "Two_Letter_Country_Code" )) 
```

```{r}
unicef <- left_join(unicef, countries_list, by = c("Code" = "Two_Letter_Country_Code" )) 
```


```{r}
names(unicef)
```


remove the unused dataframes

```{r}
rm(countries_list, countries_short)
```


```{r}
unicef %>% 
  relocate(Continent_Name, before = country_name)
```
```{r}
unicef <- unicef %>% 
  relocate(Continent_Name, before = country_name)
```



```{r}
unicef_long <- unicef %>% 
  pivot_longer(u5mr_1950:u5mr_2015, 
               names_to = "year", 
               values_to = "values")
```

```{r}
head(unicef_long)
```
Create a simplified dataset
```{r}
unicef_long %>%
  select(Continent_Name,
         country_name,
         year,
         values)
```
store

```{r}
unicef_long <- unicef_long %>%
  select(Continent_Name,
         country_name,
         year,
         values)
```


### Separate
```{r}
unicef_long %>% 
  separate(year, into = c("etc", "year"),  sep = "_")
```
```{r}
unicef_long <- unicef_long %>% 
  separate(year, into = c("etc", "year"),  sep = "_")
```

```{r}
head(unicef_long)
```

## Split combine apply

```{r}
unicef_long %>% 
  group_by(coni)
```

