---
title: "05 Recapitulation"
output:
  html_notebook: default
  html_document:
    df_print: paged
  pdf_document: default
editor_options:
  chunk_output_type: inline
---


```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE)
# Use include=TRUE for the chunks to be displayed
```

```{r}
pacman::p_load(tidyverse, # for data science
               dataMaid,  # to create data coding book
               codebook,  # alternative for codebooks
               janitor, # to clean data
               NHANES, # dataset example to play it
               gtsummary) # for tables
```

```{r}
scale_y_log10()
```


# Data entry

Best practice: separate the data entry process and keep a copy of the raw data. 
Export to CSV file and then clean it and use it for data analysis


## Dirty data




```{r}
dirty <- readr::read_csv("https://tinyurl.com/dirtyDataRSU")
```


```{r}
glimpse(dirty)
```
```{r}
View(dirty)
```

first try

```{r}
dirty %>% 
  janitor::clean_names()
```

Perfect, let's store
```{r}
cleaner <- dirty %>% 
  janitor::clean_names()
```

```{r}
glimpse(cleaner)
```

So, before

```{r}
dirty %>% 
  names()
```
After 
```{r}
cleaner %>% 
  names()
```

Duplicated data

```{r}
cleaner %>% 
  janitor::get_dupes(contains("name"))
```

```{r}
table(cleaner$subject)
```


Simple tables

One variable
```{r}
cleaner %>% 
  janitor::tabyl(subject)
```
two variables
```{r}
cleaner %>% 
  janitor::tabyl(employee_status, full_time)
```



Remove empty  rows

```{r}
cleaner <- cleaner %>% 
  janitor::remove_empty(c("rows", "cols"))
```

# NICE TABLES

## gt_summary

```{r}
NHANES %>% 
  glimpse()
```

## Create a subset dataset of NHANES

SELECT
```{r}
miniNHANES <- NHANES %>%
  select(
    Gender,
    Age,
    Race1,
    Education,
    MaritalStatus,
    HHIncomeMid,
    Weight,
    Height,
    BMI,
    BPSys1,
    BPDia1,
    DirectChol,
    TotChol,
    Diabetes,
    SleepHrsNight,
    PhysActiveDays,
    SmokeNow,
    HardDrugs
  )

```


## Make a summary table by gender
This is the Table 1. 
```{r}
miniNHANES %>% 
  select(Age, Race1, SmokeNow, Gender) %>% 
  gtsummary::tbl_summary(by = Gender) %>% 
  gtsummary::bold_labels()

```

## Histogram age by gender

```{r}
miniNHANES %>% 
  ggplot(aes(x = Age)) +
  geom_histogram(bins = 10) + 
  facet_grid(Gender~.)
```
## Point age and BMI by race and Gender

```{r}
miniNHANES %>% 
  ggplot(aes(x = Age, 
             y = BMI, 
             color = Race1)) +
  geom_point() + 
  facet_grid(Gender~Race1)
```

Add a regression line

```{r}
miniNHANES %>% 
  ggplot(aes(x = Age, 
             y = BMI, 
             color = Race1)) +
  geom_point() + 
  facet_grid(Gender~Race1) + 
  geom_smooth() # this is the regression line
```


But wait, where's the regression line?
It's hidden below all the points, so we must change the transparency of the points


```{r}
miniNHANES %>% 
  ggplot(aes(x = Age, 
             y = BMI, 
             color = Race1)) +
  geom_point(alpha = .1) +  # alpha controls the transparency
  facet_grid(Gender~Race1) + 
  geom_smooth() # this is the regression line
```
If we want a polynomial regression, we can change the equation

```{r}
miniNHANES %>% 
  ggplot(aes(x = Age, 
             y = BMI, 
             color = Race1)) +
  geom_point(alpha = .1) +  # alpha controls the transparency
  facet_grid(Gender~Race1) + 
  geom_smooth(method='lm', formula = y~poly(x,2)) # this is the regression line
```



## Create new var

hypertension

Based on syst or diast
Prevalent hypertension in National Health and Nutrition Examination Surveys (NHANES) is traditionally defined as blood pressure (BP) ≥140 mmHg systolic and/or ≥90 diastolic

```{r}
miniNHANES %>% 
  mutate()
```

## calculate mean weight and sd by race only in people over 18 years old

## Calculate age mean by race and age

## Count!

```{r}
miniNHANES %>% 
  count(Race1)
```
```{r}
miniNHANES %>%
  group_by(Gender) %>% 
  count(Race1)
```
```{r}
miniNHANES %>%
  group_by(Gender) %>% 
  count(Race1) %>% 
  pivot_wider(names_from =  Gender, values_from = n)
```

This is similar to

```{r}
miniNHANES %>% 
  select(Race1, Gender) %>% 
  gtsummary::tbl_summary(by = Gender)
```

But the advantage is that you can feed a plot with it
```{r}
miniNHANES %>%
  group_by(Diabetes, Gender) %>% 
  count(Race1) %>%
  ggplot(aes(x = Diabetes, 
             y = n, 
             fill = Gender)) + 
  geom_col()
```

nad here the same plot but dodged

```{r}
miniNHANES %>%
  group_by(Diabetes, Gender) %>% 
  count(Race1) %>%
  ggplot(aes(x = Diabetes, 
             y = n, 
             fill = Gender)) + 
  geom_bar(position="dodge", stat = "identity")
```

and here stacked
```{r}
miniNHANES %>%
  group_by(Diabetes, Gender) %>% 
  count(Race1) %>%
  ggplot(aes(x = Diabetes, 
             y = n, 
             fill = Gender)) + 
  geom_bar(position="stack", stat = "identity")
```

and stacked by 100%

```{r}
miniNHANES %>%
  group_by(Diabetes, Gender) %>% 
  count(Race1) %>%
  ggplot(aes(x = Diabetes, 
             y = n, 
             fill = Gender)) + 
  geom_bar(position="fill", stat = "identity")
```

# Codebook

```{r}
dataMaid::makeCodebook(miniNHANES)
```
Add a description for the variables

```{r}
attr(miniNHANES$Age, "shortDescription") <-
     "Age in years."
```

```{r}
attr(miniNHANES$Education, "shortDescription") <-
     "Educational level, this data was extracted from the database XXX and etc etc"
```


and now re-run the codebook


```{r}
dataMaid::makeCodebook(miniNHANES, replace=TRUE)
```


And you can make a data report of the dataset

```{r}
dataMaid::makeDataReport(miniNHANES)
```



