---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---


```{r global_options, include=FALSE}
knitr::opts_chunk$set(warning = FALSE,
                      message = FALSE)
# Use include=TRUE for the chunks to be displayed
```




If you have some issues, please uncomment the line that install the package, removing the # at the beginning

# Packages
```{r message=FALSE, warning=FALSE, echo=T}
# install.packages("pacman") # this is a package that allows to manage other packages
library(pacman)
```


```{r, message=FALSE, echo=FALSE, results='hide', warning=FALSE}
pacman::p_load(tidyverse, 
               gtsummary, 
               janitor, 
               visdat, 
               ggpubr, 
               naniar, 
               NHANES, 
               palmerpenguins, 
               gapminder, 
               coronavirus)
```

# Review

Shortcuts

CTRL + ALT + I


```{r}
x <- c(1:55)
```

```{r}
x %>% summary
```

# Explore the dataset

```{r}
penguins
```
```{r}
dim(penguins)
```
```{r}
glimpse(penguins)
```
```{r}
penguins %>% summary()
```
```{r}
class(penguins$year)
```
```{r}
penguins$year <- as_factor(penguins$year)
```

```{r}
class(penguins$year)
```

```{r}
View(penguins)
```




# Distributions


```{r}
penguins %>% 
  ggplot(aes(x = body_mass_g)) +
  geom_histogram(bins = 10) +
  aes(fill = species) +
  facet_wrap(~species)
```



```{r}
penguins %>% 
  
  drop_na() %>% 
  
  ggplot(aes(x = body_mass_g)) +
  geom_histogram(bins = 10) +
  aes(fill = species) +
  facet_grid(sex ~ species)
```







# Associations

x = cont 
y = cont

```{r}
penguins %>% 
  ggplot(aes(x = bill_length_mm, 
             y = bill_depth_mm)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  aes(color = species)
```


# Comparison

is the weight similar between penguins?

body_mass
species factor

```{r}
penguins %>% 
  ggplot(aes(x = species, 
             y = body_mass_g)) +
  geom_boxplot()
```

Compare the bill_length across all species

```{r}
penguins %>%
  drop_na() %>% 
  ggplot(aes(x = species,
             y = bill_length_mm)) +
  geom_boxplot() +
  aes(color = sex)
```


-------------------------------------------

```{r}
NHANES
```

```{r}
View(NHANES)
```

```{r}
glimpse(NHANES)
```

```{r}
table(NHANES$Gender)
```

```{r}
table(NHANES$Race1)
```

```{r}
table(NHANES$Gender, 
      NHANES$Race1)
```
```{r}
table(NHANES$SurveyYr)
```

```{r}
NHANES %>% 
  ggplot(aes(x = Age)) + 
  geom_histogram(bins = 15) +
  facet_grid(Gender ~ . )
```

```{r}
NHANES %>% 
  ggplot(aes(x = Height, 
             y = Weight)) + 
  geom_point(size = .01)
```
Is there any association between the Head Circunference and the income?

```{r}
NHANES %>% 
  ggplot(aes(x = HeadCirc, 
             y = HHIncomeMid)) + 
  geom_point() + 
  geom_smooth(method = "lm") +
  aes(color = Gender)
```




# Trends




