---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
editor_options:
  chunk_output_type: inline
---

# packges

```{r}
pacman::p_load(tidyverse, 
               gapminder, 
               palmerpenguins, 
               janitor, 
               gtsummary, 
               visdat)
```


# Recapitulation

# EDA

```{r}
head(penguins)
```

```{r}
dim(penguins)
```
```{r}
summary(penguins)
```
```{r}
visdat::vis_dat(penguins)
```

### Data visualizations

#### Points
bill_length_mm
vs 
bill_depth_mm

```{r}
penguins %>% 
  ggplot(aes(x = bill_length_mm, 
             y = bill_depth_mm)) +
  geom_point()
```

```{r}
penguins %>% 
  ggplot(aes(x = bill_length_mm, 
             y = bill_depth_mm)) +
  geom_point() +
  aes(color = sex)
```

```{r}
penguins %>% # my date
  drop_na() %>%  # i will dropn the NA values
  ggplot(aes(x = bill_length_mm, 
             y = bill_depth_mm)) +
  geom_point() + 
  aes(color = sex)
```

##### Facetting

```{r}
penguins %>% # my date
  drop_na() %>%  # i will dropn the NA values
  ggplot(aes(x = bill_length_mm, 
             y = bill_depth_mm)) +
  geom_point() + 
  aes(color = sex) +
  facet_wrap(~species)
```
```{r}
penguins %>% # my date
  drop_na() %>%  # i will drop the NA values
  ggplot(aes(x = bill_length_mm, 
             y = bill_depth_mm)) +
  geom_point() + 
  aes(color = sex) +
  facet_grid(island ~ species)
```

### One variable continuous
make an histogram with body_mass_g


```{r}
penguins %>%
  drop_na() %>% 
  ggplot(aes(x = body_mass_g)) +
  geom_histogram(bins = 12) +
  facet_grid(species ~ sex) +
  theme_minimal()
```
```{r}
pacman::p_load(ggthemes)
```

```{r}
penguins %>%
  drop_na() %>% 
  ggplot(aes(x = body_mass_g)) +
  geom_histogram(bins = 12) +
  facet_grid(species ~ sex) +
  ggpubr::theme_pubclean()
```


### Two variables continuous

```{r}
penguins %>%
  drop_na() %>%
  ggplot(aes(x = body_mass_g,
             y = bill_depth_mm)) +
  geom_point() +
  aes(color = sex) +
  facet_grid(sex ~ species) +
  theme_minimal() +
  labs(
    title = "Bill vs Weight", 
    y = "Bil Depth mm", 
    x = "Body mass g", 
    color = "Sex"
  )
```
#### Explore correlations

```{r}
penguins %>%
  drop_na() %>%
  ggplot(aes(x = body_mass_g,
             y = bill_depth_mm)) +
  geom_point() +
  aes(color = sex) +
  theme_minimal() +
  labs(
    title = "Bill vs Weight", 
    y = "Bil Depth mm", 
    x = "Body mass g", 
    color = "Sex"
  ) + 
  geom_smooth()
```
### Two variables continuous vs nominal

body_mass_g by species

```{r}
penguins %>% 
  drop_na() %>% 
  ggplot(aes(x = species,
             y = body_mass_g)) +
  geom_boxplot() +
  aes(color = sex)
```
### Trends

```{r}
gapminder
```
```{r}
gapminder <- gapminder::gapminder
```

# Data wrangling

## Filter

```{r}
gapminder %>% 
  filter(country == "Germany"  ) 
```
```{r}
gapminder %>% 
  filter(continent == "Europe" & year == 2007)
```

Asia and 1952
```{r}
gapminder %>% 
  filter(continent=="Asia" & year == "1952")
```
```{r}
gapminder %>% 
  filter(continent %in% c("Asia", "Europe") & year %in% c(1952, 2002) ) 
```


```{r}
selected_countries <- c("Bahrain", "Albania")
```



## Select
```{r}
gapminder %>% 
  select(continent, 
         lifeExp)
```
```{r}
gapminder %>%
  filter(country %in% selected_countries)
```



## Mutate

1 United States Dollar equals 0.84 Euro

```{r}
gapminder %>% 
  mutate(gpd_in_euros = gdpPercap * 0.84)
```



## Group

```{r}
gapminder %>% 
  group_by(continent, year)
```


## Summarise
```{r}
gapminder %>% 
  mutate(gpd_in_euros = gdpPercap * 0.84) %>% 
  group_by(continent, year) %>% 
  summarise(meanGdpperCapEur = mean(gpd_in_euros))
```
```{r}
gapminder %>% 
  mutate(gpd_in_euros = gdpPercap * 0.84) %>% 
  group_by(continent, year) %>% 
  summarise(meanGdpperCapEur = mean(gpd_in_euros)) %>% 
  ggplot(aes(x = year, 
             y = meanGdpperCapEur, 
             color = continent)) + 
  geom_line() +  
  scale_y_log10()
 
```

```{r}
penguins %>% 
  drop_na() %>% 
  select(body_mass_g, bill_length_mm, species, island) %>% 
  gtsummary::tbl_summary(by = species) %>% 
  gtsummary::bold_labels()
```






## Pivot



```{r}
unicef <- read_csv("https://bit.ly/unicef-wide")
```


```{r}
head(unicef)
```

```{r}
unicef <- unicef %>% 
  janitor::clean_names()
```



### Longer

```{r}
unicef %>% 
  pivot_longer(u5mr_1950:u5mr_2015, 
               names_to = "year", 
               values_to = "value")
```
```{r}
unicef_long <- unicef %>% 
  pivot_longer(u5mr_1950:u5mr_2015, 
               names_to = "year", 
               values_to = "value")
```

```{r}
unicef_long
```

```{r}
unicef_long <- unicef_long %>%
  separate(year,
           into = c("etc", "year"),
           sep = "_") %>% 
  select(-etc)
```


```{r}
head(unicef_long)
```

### Wider
```{r}
gapminder %>% 
  group_by(continent, year) %>% 
  summarise(LifeExpMean = mean(lifeExp))
```
```{r}
gapminder %>% 
  group_by(continent, year) %>% 
  summarise(LifeExpMean = mean(lifeExp)) %>% 
  pivot_wider(names_from = year, 
              values_from = LifeExpMean )
```

## Join dataset

```{r}
unicef
```


```{r}
countries_short <- read_csv("https://pkgstore.datahub.io/core/country-list/data_csv/data/d7c9d7cfb42cb69f4422dec222dbbaa8/data_csv.csv")


countries_list <- read_csv("https://datahub.io/JohnSnowLabs/country-and-continent-codes-list/r/country-and-continent-codes-list-csv.csv")
```
```{r}
head(countries_short)
```


```{r}
countries_list
```



```{r}
unicef
```

```{r}
left_join(unicef, 
          countries_short, 
          by = c("country_name" = "Name"))
```
```{r}
unicef <- left_join(unicef, 
          countries_short, 
          by = c("country_name" = "Name"))
```

```{r}
left_join(unicef, 
          countries_list, 
          by = c("Code" = "Two_Letter_Country_Code"))
```

```{r}
unicef <- left_join(unicef, 
          countries_list, 
          by = c("Code" = "Two_Letter_Country_Code"))
```


```{r}
names(unicef)
```
calculate the mortality rate by continent and year

```{r}
unicef
```


What is the trend in the mortality rate by continent?

```{r}
unicef %>% 
  # first reformat from wide to long
  pivot_longer(u5mr_1950:u5mr_2015, 
               names_to = "year", 
               values_to = "value") %>% 
  # select only relevant columns 
  select(continent_name, year, value) %>% 
  # separate the year column 
  separate(year, into = c("delete", "year"), 
           sep = "_") %>% 
  # now delete the delete column 
  select(-delete) %>% 
  # group and summarize
  group_by(continent_name, year) %>% 
  summarise(meanMrtRate5y = mean(value)) %>% 
  ggplot(aes(x = year, 
             y = meanMrtRate5y, 
             color = continent_name)) + 
  geom_point()
```































