---
title: COVID data
author: Sergio Uribe
date: March 22, 2005
output:
  html_document:
    toc_float: true
    theme: cerulean
    toc: true
---


```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
# Use include=TRUE for the chunks to be displayed
```

 
# Packages
```{r}
# if (!require("pacman")) install.packages("pacman") # run in a new installation
pacman::p_load(tidyverse,  # for data science
               janitor,    # for data cleaning 
               gt,         # for tables
               ggpubr,     # for graphs publication themes
               gtsummary,  # for data summaries
            #  epitool,    # for Epi
            #  pubh,       # for public health
               finalfit,   # for model reporting
               naniar,     # for NAs management
               cowplot,    # for various graphs 
               santoku,    # to create new vars 
               ggsignif,   # to add sign values to ggplots
               GGally      # for regression diagnostic plots
               )
```

Graph theme
```{r}
# theme_set(theme_minimal())
```


# Dataset
```{r}
df <- read_csv("https://opendata.ecdc.europa.eu/covid19/casedistribution/csv")
```
```{r}
glimpse(df)
```




# EDA

```{r}
summary(df)
```
```{r}
table(df$continentExp)
```
```{r}
df %>% 
  filter(continentExp == "Other")
```



------------------

Use and delete

## gtsummary
trial2 %>%
  # build base summary table
  tbl_summary(
    # split table by treatment variable
    by = trt,
    # change variable labels
    label = list(marker ~ "Marker, ng/mL",
                 stage ~ "Clinical T Stage"),
    # change statistics printed in table
    statistic = list(all_continuous() ~ "{mean} ({sd})",
                     all_categorical() ~ "{n} / {N} ({p}%)"),
    digits = list("marker" ~ c(1, 2))
  ) %>%
  # add p-values, report t-test, round large pvalues to two decimal place
  add_p(test = list(marker ~ "t.test"),
                 pvalue_fun = function(x) style_pvalue(x, digits = 2)) %>%
  # add statistic labels
  add_stat_label() %>%
  # bold variable labels, italicize levels
  bold_labels() %>%
  italicize_levels() %>%
  # bold p-values under a given threshold (default is 0.05)
  bold_p(t = 0.2) %>%
  # include percent in headers
  modify_header(stat_by = "**{level}**, N = {n} ({style_percent(p, symbol = TRUE)})")
  

## Locate and change NAs
which(is.na(col))
df$col[is.na(df$col)] <- new_value

## Remove columns with certain level NA
x <- df[ lapply( df, function(x) sum(is.na(x)) / length(x) ) < 0.4 ]

## Search and replace 
df <- dplyr::mutate_if(tibble::as_tibble(df), 
                                is.character, 
                                stringr::str_replace_all, 
                       pattern = "oldVar",replacement = "newVar")
                       

