---
title: "Datasets"
output: html_notebook
---

First install the package that allow to manage others packages

Write, inside a chunk (Insert \> Code Chunk \> R or CTRL + ALT + I)

install.packages("pacman")

```{r}
pacman::p_load(tidyverse, # several datascience packages
               gtsummary) # for tables) 
```

## Datasets

```{r}
pacman::p_load(palmerpenguins)
```

```{r}
head(penguins)
```

```{r}
pacman::p_load(babynames)
```

```{r}
head(babynames)
```

```{r}
head(mtcars)
```

```{r}
pacman::p_load(gapminder)
```

```{r}
head(gapminder)
```

```{r}
pacman::p_load(coronavirus)
```

```{r}
dim(coronavirus)
```

```{r}
names(coronavirus)
```

```{r}
coronavirus %>% 
  filter(country == "Latvia")
```

```{r}
head(esoph)
```

A data frame with records for 88 age/alcohol/tobacco combinations.

| \[,1\] | "agegp"     | Age group           | 1 25--34 years |
|-------:|:------------|:--------------------|:---------------|
|        |             |                     | 2 35--44       |
|        |             |                     | 3 45--54       |
|        |             |                     | 4 55--64       |
|        |             |                     | 5 65--74       |
|        |             |                     | 6 75+          |
| \[,2\] | "alcgp"     | Alcohol consumption | 1 0--39 gm/day |
|        |             |                     | 2 40--79       |
|        |             |                     | 3 80--119      |
|        |             |                     | 4 120+         |
| \[,3\] | "tobgp"     | Tobacco consumption | 1 0-- 9 gm/day |
|        |             |                     | 2 10--19       |
|        |             |                     | 3 20--29       |
|        |             |                     | 4 30+          |
| \[,4\] | "ncases"    | Number of cases     |                |
| \[,5\] | "ncontrols" | Number of controls  |                |

```{r}
head(HairEyeColor)
```

```{r}
head(InsectSprays)
```

```{r}
InsectSprays %>% 
  ggplot(aes(x = count)) +
  geom_histogram(bins = 8) + 
  facet_grid(spray ~. )
```

```{r}
head(morley)
```

```{r}
morley %>% 
  ggplot(aes(x = Speed)) + 
  geom_histogram(bins = 10) + 
  facet_grid(Run ~ Expt)
```

```{r}
summary(aov(Speed ~ Expt * Run, data =  morley))
  
```

```{r}
morley %>% 
  ggplot(aes(x = as.factor(Expt), 
             y = Speed)) + 
  geom_boxplot() + 
  geom_jitter(alpha = .3)
```

```{r}
head(sleep)
```

A data frame with 20 observations on 3 variables.

| \[, 1\] | extra | numeric | increase in hours of sleep |
|--------:|:------|:--------|:---------------------------|
| \[, 2\] | group | factor  | drug given                 |
| \[, 3\] | ID    | factor  | patient ID                 |

```{r}
sleep %>% 
  ggplot(aes(x = group, 
             y = extra)) + 
  geom_violin() + 
  geom_jitter(alpha = .3, width = .1, color = "red") + 
  geom_boxplot(alpha = .1, width = .25, color = "light blue")
```

## fivethirtyeightdata 

```{r}
pacman::p_load(fivethirtyeight)
```

```{r}
head(classic_rock_raw_data)
```

## trick

The chickweight dataset is the best.

Don't trust me? There's something horribly wrong with that dataset. But it's hidden! Can you find it? Hint; you'll need to combine ggplot with dplyr in order to understand the pattern in it.

```{r}
pacman::p_load(ChickWeight)
```

```{r}
head(ChickWeight)
```

```{r}
summary(ChickWeight)
```

```{r}
pacman::p_load(NHANES)
```

```{r}
head(NHANES)
```

```{r}
head(starwars)
```

```{r}
starwars %>% 
  group_by(species) %>% 
  summarise(mean_mass = mean(mass)) %>% 
  drop_na() %>% 
  arrange(desc(mean_mass)) %>% 
  ggplot(aes(x = fct_reorder(species, mean_mass), 
             y = mean_mass)) + 
  geom_col() +
  coord_flip() + 
  scale_y_log10()
```
