---
title: "Introduction to Exploratory Data Analysis"
output: html_notebook
---

# Packages

```{r}
pacman::p_load(tidyverse, 
               NHANES, 
               janitor, 
               visdat, 
               naniar, 
               psych, # for describe
               sjPlot)
```

# Dataset

```{r}
data("NHANES")
```

# Explore the dataset

```{r}
head(NHANES)
```

```{r}
tail(NHANES)
```

```{r}
colnames(NHANES)
```

```{r}
dim(NHANES)
```

## Missing values

```{r}
NHANES %>% 
  naniar::miss_var_summary() %>% 
  filter(pct_miss > 65)
```

With visdat package

```{r}
NHANES %>% 
  visdat::vis_dat()
```

```{r}
naniar::gg_miss_var(NHANES)
```

```{r}
NHANES %>% 
  gg_miss_upset()
```

Could be any influence of the missing data on certain variables?

```{r}
NHANES %>%
  bind_shadow() %>%
  ggplot(aes(x = Age,
             fill = Poverty_NA)) + 
  geom_density(alpha = 0.5)
```

```{r}
NHANES %>%
  select(-contains(c("Preg", "Bab"))) %>% 
  group_by(Gender) %>%
  miss_var_summary() %>% 
  filter(pct_miss > 65)
```

## Data exploration

```{r}
str(NHANES)
```

Compare with glimpse

```{r}
glimpse(NHANES)
```

```{r}
summary(NHANES)
```

```{r}
summary(NHANES$Age)
```

```{r}
psych::describe(NHANES)
```

# Summarizing values

1.  How many variables to manipulate

-   1A. single variable

-   1B. more than a variable

2.  How many functions to apply to each variable

-   2A. single function

-   2B. more than one function

resulting in the following four cases:

-   Case 1: apply one function to one variable

-   Case 2: apply many functions to one variable

-   Case 3: apply one function to many variables

-   Case 4: apply many functions to many variables

### **Case 1: apply one function to one variable**

```{r}
NHANES %>% 
  summarise(mean_age = mean(Age))
```

or by gender

```{r}
NHANES %>% 
  group_by(Gender) %>% 
  summarise(mean_age = mean(Age))
```

### **Case 2: apply many functions to one variable**

```{r}
NHANES %>% 
  summarise(min_age = min(Age), max_age = max(Age))
```

By gender

Note the na.rm = T

```{r}
NHANES %>% 
  group_by(Gender) %>% 
  summarise(min_age = min(Pulse, na.rm = T), 
            max_age = max(Pulse))
```

or with summarize_each

```{r}
NHANES %>% 
  summarise(across("Age",  
                   list(min = min, Max = max)))
```

grouping

```{r}
NHANES %>% 
  group_by(Gender) %>% 
  summarise(across(c(Age), 
                   ~ mean(.x, na.rm = TRUE)))
```

Change the name of columns

```{r}
NHANES %>% 
  group_by(Gender) %>% 
  summarise(across(c(Age), 
                   ~ mean(.x, na.rm = TRUE), 
                   .names = "mean_{.col}"))
```

### **Case 3: apply one function to many variables**

```{r}
NHANES %>% 
  summarise(mean_age = mean(Age), 
            mean_pulse = mean(Pulse, na.rm = T))
```

grouping by gender

```{r}
NHANES %>% 
  group_by(Gender) %>% 
  summarise(mean_age = mean(Age), 
            mean_pulse = mean(Pulse, na.rm = T))
```

```{r}
NHANES %>% 
  summarise(across(c("Age", "Pulse"), 
                   ~mean(.x, na.rm = T)))
  
```

```{r}
NHANES %>% 
  group_by(Gender) %>% 
  summarise(across(c("Age", "Pulse"), 
                   ~mean(.x, na.rm = T)))
```

### **Case 4: apply many functions to many variables**

```{r}
NHANES %>% 
  summarise(min_Age = min(Age) , 
            min_Pulse = min(Pulse, na.rm = T), 
            max_Age = max(Age) , 
            max_Pulse = max(Pulse, na.rm = T))
```

```{r}
NHANES %>% 
  summarise(across(c("Age", "Pulse"), 
                    list(mean, sd ), na.rm = T))
```

```{r}
NHANES %>% 
  group_by(Gender) %>% 
  summarise(across(c("Age", "Pulse"), 
                    list(mean, sd),
                   # .names = "mean_{.col}", 
                   na.rm = T))
```

## Bonus, use significant numbers

```{r}
# Packages
require(NHANES, # the dataset
        dplyr, 
        magritte)

# the dataset
data(NHANES)


# example, no errors
NHANES %>% 
  group_by(Gender) %>% 
  summarise(across(c("Age", "Pulse"), 
                    list(mean, min, max), na.rm = T)) 


```

```{r}
# example, errors, some columns are CHR now
NHANES %>% 
  group_by(Gender) %>% 
  summarise(across(c("Age", "Pulse"), 
                    list(mean, min, max), na.rm = T)) %>% 
  mutate(across(is.numeric, format, 1))
```
