---
title: "Lasma dataset"
output:
  html_notebook: default
  pdf_document: default
  word_document: default
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  message = F, 
  warning = F
)
```


# Load some packages

```{r}

pacman::p_load(tidyverse, 
               sjPlot)


```
Load the dataset
```{r}
df <- read_csv("https://j.mp/2JJYkeN")
```

Explore the dataset

```{r}
head(df)
```

# Exploratory data analysis

```{r}
glimpse(df)
```
Some thoughts: 
Year should be integer or factor


Check the amount of NA data

For this I will use the visdat package

```{r}
pacman::p_load(visdat)
```

and plot
```{r}
df %>% 
  visdat::vis_dat()
```
Lot of NAs...

Let's check some variables

```{r}
table(df$Gender)
```
```{r}
table(df$`Power position`)
```

```{r}
table(df$Gender, df$`Power position`)
```
Seems to me that the data are divided in 

_demographic data_

 [1] "RespondentID"           "Power position"         "Participation"          "Year"                   "Gender"                
  [6] "Language"               "Parties"                "YearsMP"                "Chairman"               "Specialisation"        
 [11] "8.LETA/BNS"    
 
_Some questions about use of social media_

"Monitoring"             "Oft_radio_TV"           "Oft_newspaper"          "Oft_publication"       
 [16] "FB"                     "draugiem.lv"            "Instagram"              "vkontakte"              "Odnoklassniki"         
 [21] "Twitter"                "dont_use" 
 
 _some perceptions about their utility_
 
 
 the rest of the columns
 
 
 We can use an special package to create summary table ready for publication, it's called gtsummary
 
```{r}
pacman::p_load(gtsummary)
```

Learn about the package here: https://github.com/ddsjoberg/gtsummary

For example, a table of the demographic data by power position would be: 


```{r}
df %>% 
  select(`Power position`:Specialisation) %>%  # here I select the desired columns to summarize
  gtsummary::tbl_summary(by = `Power position`)
```
As you can see, there is _a lot_ of data to clean here. 

I _strongly suggest_ you to learn two of the tidyverse packages: 

https://stringr.tidyverse.org/

https://forcats.tidyverse.org/



# Questions

What are MPs perception on media power compared to other actors (Columns Agenda_TV_radio and Agenda_written_press compared to Agenda_prime; Agenda_ministers; Agenda_MP; Agenda_parties; Agenda_interest_groups). The survey was in Likert scale 1 - 5);

I don't have much experience with this, so I google: 
https://www.google.com/search?q=lickert+data+r&oq=lickert+data+r&aqs=chrome..69i57.7470j0j1&sourceid=chrome&ie=UTF-8


I will subset the dataset, selecting some columns only

```{r}
mydf <- df %>% 
  select(Journalists_indep:Agenda_written_press)
```


after googling some _hours_ found some packages that can help to analyze Likert data

```{r}

mydf %>% 
  sjPlot::plot_likert() 
```
Please, read the documentation and options of the package here: 
https://cran.r-project.org/web/packages/sjPlot/sjPlot.pdf


Also you can order by some options, see page 25, the option sort.frq = from the plot_likert function

```{r}
mydf %>%
  sjPlot::plot_likert(sort.frq = "pos.asc" )  # see page 25
```
and you can create different plot for different groups, for example for men and women

```{r}
df %>%
  filter(Gender == "Man") %>% # filter only men
  select(Agenda_ministers:Agenda_written_press) %>% # and these range of columns
  sjPlot::plot_likert()  # see page 25
```


```{r}
df %>%
  filter(Gender == "Woman") %>% # filter only men
  select(Agenda_ministers:Agenda_written_press) %>% # and these range of columns
  sjPlot::plot_likert()  # see page 25
```


:)






# Learn more

learn about the sjPlot  package here: 

https://strengejacke.github.io/sjPlot/



# Learn even more!

Check the survey package, specially designed to analyze data from surveys

http://r-survey.r-forge.r-project.org/survey/

since is quite an old package, there is a tool that allows you to work with dplyr from tidyverse package: 

https://github.com/gergness/srvyr





