---
title: "title"
author: '[Sergio Uribe](https://sergiouri.be)'
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: no
    code_folding: hide
---

change in in /usr/lib/rstudio/resources/templates

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE, out.width="75%")
# Use include=TRUE for the chunks to be displayed
```

# Packages

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# if (!require("pacman")) install.packages("pacman") # run in a new installation
pacman::p_load(tidyverse, 
               gapminder, 
               nycflights13
)
```

# Dataset

```{r}
gapminder
```

```{r}
theme_set(theme_minimal())
```

## Life expectancy in the Americas in 2007

```{r}
gapminder %>%
  filter(year == 2007,
         continent == "Americas") %>%
  ggplot(aes(x = country, 
             y = lifeExp)) +
  geom_point() + 
  coord_flip()
```

Default order is alphabetic, from bottom to top

```{r}
gapminder %>%
  filter(year == 2007,
         continent == "Americas") %>%
  ggplot(aes(x = fct_reorder(country, lifeExp), 
             y = lifeExp)) +
  geom_point() + 
  coord_flip()
```

## We can also order facets

```{r}
gapminder %>%
  filter(country %in% c("Norway", "Portugal", "Spain", "Austria")) %>%
  ggplot(aes(x = year, 
             y = lifeExp)) + 
  geom_line() +
  facet_wrap( vars(country), nrow = 1)

```

-   Default ordering is alphabetic; there's no good reason for this ordering

    ```{r}
    gapminder %>%
      filter(country %in% c("Norway", "Portugal", "Spain", "Austria")) %>%
      mutate(country = fct_reorder(country, lifeExp)) %>% 
      ggplot(aes(x = year,  
                 y = lifeExp)) + 
      geom_line() +
      facet_wrap( vars(country), nrow = 1)
    ```

Alternative orderings: By smallest value per facet

```{r}

gapminder %>%
  filter(country %in% c("Norway", "Portugal", "Spain", "Austria")) %>%
  mutate(country = fct_reorder(country, lifeExp, min)) %>% # order by minimum
  ggplot(aes(year, lifeExp)) + geom_line() +
  facet_wrap(vars(country), nrow = 1)
```

By the smallest difference

```{r}
gapminder %>%
  filter(country %in% c("Norway", "Portugal", "Spain", "Austria")) %>%
  # order by custom function: here, difference between max and min
  mutate(country = fct_reorder(country, lifeExp, 
                               function(x) {max(x) - min(x)
                                 })) %>%
  ggplot(aes(year, lifeExp)) + geom_line() +
  facet_wrap(vars(country), nrow = 1)
```

```{r}

flight_data <- flights %>% # take data on individual flights
  left_join(airlines) %>%  # add in full-length airline names
  select(name, carrier, flight, year, month, day, origin, dest) # pick columns of interest
```

```{r}
flight_data
```

## Flights out of New York City in 2013

```{r}
flight_data %>%
  ggplot(aes(y = name)) + 
  geom_bar()
```

```{r}
flight_data %>%
  mutate(
    name = fct_infreq(name)
  ) %>%
  ggplot(aes(y = fct_rev(name))) + 
  geom_bar()
```

```{r}
flight_data %>%
  mutate(# order after lumping
    name = fct_infreq(fct_lump(name, 7))) %>%
  ggplot(aes(y = fct_rev(name))) +
  geom_bar()
```

```{r}

flight_data %>%
  mutate(# order before lumping
    name = fct_lump(fct_infreq(name), 7)) %>%
  ggplot(aes(y = fct_rev(name))) +
  geom_bar()
```

```{r}
flight_data %>%
  mutate(
    name = fct_lump(fct_infreq(name), 7),
    # Use `fct_other()` to manually lump all
    # levels not called "Other" into "Named"
    highlight = fct_other(name,
                          keep = "Other", 
                          other_level = "Named")
  ) %>%
  ggplot() +
  aes(y = fct_rev(name),
      fill = highlight) +
  geom_bar()
```
```{r}
flight_data %>%
  mutate(
    name = fct_lump(fct_infreq(name), 7),
    # Use `fct_other()` to manually lump all
    # levels not called "Other" into "Named"
    highlight = fct_other(name,
                          keep = "Other", other_level = "Named")
  ) %>%
  ggplot() +
  aes(y = fct_rev(name),
      # reverse fill aesthetic
      fill = fct_rev(highlight)) +
  geom_bar()
```

Summary of key factor manipulation functions
Function	Use case	Documentation
fct_relevel()	Change order of factor levels manually	click here
fct_infreq()	Put levels in descending order of how frequently each level occurs in the data	click here
fct_rev()	Reverse the order of factor levels	click here
fct_reorder()	Put levels in ascending order determined by a numeric variable or summary function	click here
fct_lump_n()	Retain the n most frequent levels and lump all others into "Other"	click here
fct_other()	Manually group some factor levels into "Other"	click here
For more options, check out the reference documentation of the forcats package