---
title: "First class"
output: html_notebook
---


```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  include = T,
  # eval = T,
  include = T,
  echo = T
)
```



# Packages

The pacman command allows to easy install other packages


```{r}
pacman::p_load(tidyverse, 
               palmerpenguins)
```


# Dataset
Now I will load the penguins dataset to my enviroment (check the upper right panel)

```{r}
data(penguins)
```


# Explore


With the View() command you can open the dataset in an external tab, as an spreadsheet, but you will not be able to modify it. And that's a good thing, remember: don't touch your data!

```{r}
View(penguins)
```


With the dim() commando, you can see the DIMensions of the dataset, in rows and columns

```{r}
dim(penguins)
```

I can explore the first six rows of the dataset, and this one of the first things that I do when I have some data
```{r}
head(penguins)
```

and also the last six rows
```{r}
tail(penguins)
```
With the str(), you can see the STRucture of your dataset (or dataframe). 
Note that indicates the **class** and the **levels** of each categorical (or **factor**) variable.
Also, you see the dimensions in the first line. 

It tell us that this dataset is a tibble, a especial kind of dataset, that has 344 rows per 8 columns

```{r}
str(penguins)
```
The str() is *super* useful and also one of the first things that you have to examine about your data, and consult several times during your data analysis.


The *tidyverse* version of str() is the glimpse() commando

```{r}
glimpse(penguins)
```

If you modify the width of this panel, making it very 


```{r}
summary(penguins)
```

```{r}
class(penguins$year)
```
```{r}
penguins$year <- as.factor(penguins$year)
```

now the year variable should be correctly formatted

```{r}
summary(penguins$year)
```
Now it's fixed!


# Creating figures with ggplot

ggplot stands for grammar of graphics. 

Just as language has rules with which to communicate any idea, graphics have a grammar. If we understand the components of a graph and how the different parts that make up a graph are related, we can create any type of visualization to communicate a result. 

Essentially, to make a graphic we need three things: 
1. data
2. map the data with some aesthetics, for example, in x, y axes, with colors, shapes or sizes, and
3. a geometric shape that represents the data in the mapping 


So the first thing we need is the data

```{r}
penguins %>%  # the data
  ggplot()    # we tell: hey ggplot, take this data, and wait for instructions
```

Secondly, we map the variables of interest

Usually this is done this way: 

```{r}
penguins %>% 
  ggplot(mapping = aes(x = bill_length_mm,  #hey ggplot, map these aesthetic variables, in the x axis this variable
                       y = body_mass_g)) # and in the y axis this other
  
```


I personally omit the mapping, so I write

```{r}
penguins %>% 
  ggplot(aes(x = bill_length_mm, 
             y = body_mass_g))
```

and is the same result . 

Ahora deberíamos decirle a ggplot cómo mapear esas variables x e y. En este caso, elegiremos mapearlas como puntos.  

```{r, include=TRUE}
penguins %>% 
  ggplot(aes(x = bill_length_mm, 
             y = body_mass_g)) + 
  geom_point()  # this is the geometric object that will map the aesthetic
```


Later, we add more layers as aesthetics elements. 

For example, we can add a color to each point. We add this as an additional aesthetic, see code below

```{r}
penguins %>%
  drop_na() %>%  
  ggplot(aes(x = bill_length_mm, 
             y = flipper_length_mm)) +
  geom_point() +
  aes(color = species)  # this is the additional layer
```

or we can add as an additional aesthetic with the former aesthetics variables used
```{r}
penguins %>%
  ggplot(aes(x = bill_length_mm,
             y = flipper_length_mm,
             color = species))  +   # in this case, here
  geom_point()
```
And is the same result. 
Now take a moment to look at the graph and reflect on the information provided by the data


## EXERCISES




## Tuning your graph

### Themes
You can change the theme of the graph, the visual appearence, changing the layer that control this, with the commando *theme* and choosing your theme of preference

```{r}
penguins %>%
  ggplot(aes(x = bill_length_mm,
             y = flipper_length_mm,
             color = species))  +   
  geom_point() + 
  theme_minimal()   # this is one example
```
```{r}
penguins %>%
  ggplot(aes(x = bill_length_mm,
             y = flipper_length_mm,
             color = species))  +   
  geom_point() + 
  theme_dark() # now a dark theme
```
Try others!

try
theme_linedraw()
theme_light()

```{r}
penguins %>%
  ggplot(aes(x = bill_length_mm,
             y = flipper_length_mm,
             color = species))  +   # in this case, here
  geom_point() 
```
You can also install more themes or [create your own theme](https://emanuelaf.github.io/own-ggplot-theme.html)

Some packages with additional themes are: 

[ggpubr](http://www.sthda.com/english/articles/24-ggpubr-publication-ready-plots/) (I use mostly the minimal and some of the ggpubr package)
and [ggthemes](https://rpubs.com/Mentors_Ubiqum/ggthemes_1)


To install use
```{r}
pacman::p_load(ggthemes)
```

and then
```{r}
penguins %>%
  ggplot(aes(x = bill_length_mm,
             y = flipper_length_mm,
             color = species))  +   # in this case, here
  geom_point() + 
  ggthemes::theme_economist()  # here I say" from the package ggthemes use the theme_economist()
```
You can try these themes:   
 - ggthemes::theme_solarized()
 - ggthemes::theme_excel()



# VOY AQUI

### Headings and labels



# Note

You will find this kind of notation, where the data goes inside the ggplot command. For example: 

```{r}
ggplot(data = penguins) +
  (mapping = aes(x = species, 
                 y = body_mass_g, 
                 color = sex)) + 
  geom_boxplot() + 
  theme_minimal()
```
I prefer to leave the data out of the ggplot command, since it's easier to perform some data transformation and then plot it. 

For example, in 


































