---
title: "02 Session: Data Visualizations"
author: Sergio Uribe
output: html_notebook
editor_options: 
  chunk_output_type: inline
---



This code at the beginning tells all others not to show warnings or messages


```{r global_options, include=FALSE}
knitr::opts_chunk$set(warning = FALSE,
                      message = FALSE, 
                      echo = TRUE)
# Use include=TRUE for the chunks to be displayed
```

# Review

Shortcuts

CTRL + ALT + I

ALT + -

```{r}
x <- c(1:50)
```


CTRL + ALT + I

```{r}
x %>% 
  summary
```

Beautify your code

CTRL + SHIFT + A

```{r}
x<-c(1:53) 
```



# Packages
```{r}
# install.packages("pacman") # this is a package that allows to manage other packages
library(pacman)
```

```{r}
pacman::p_load(tidyverse, 
               gtsummary, 
               janitor, 
               visdat, 
               ggpubr, 
               naniar, 
               NHANES, 
               palmerpenguins, 
               gapminder, 
               coronavirus)
```


# Explore the dataset

```{r}
penguins
```

```{r}
View(penguins)
```

```{r}
dim(penguins)
```

```{r}
head(penguins)
```

```{r}
tail(penguins)
```

```{r}
summary(penguins)
```

To access one specific variable, use the dollar sign

```{r}
summary(penguins$bill_depth_mm)
```



Check the class of the variables
```{r}
class(penguins$year)
```

# Types of variables in R

```{r}
glimpse(penguins)
```
_fct_ : factors, can be nominal with two levels (as male, female) or more categories. By default R try to order the levels, from 1 and on or A and on. 

_int_ : integer numbers
_dbl_ : any number, even with decimals

Other types
_logical_
_date_ 

You can change the class of a variable

```{r}
class(penguins$year)
```

```{r}
summary(penguins$year)
```

Useless, so we can change the variable

```{r}
penguins$year <- as_factor(penguins$year)
```

```{r}
class(penguins$year)
```

```{r}
summary(penguins$year)
```
# Data visualizations

## The grammar of graphics

Three things: 
 1. Data
 1. the mapping of the aesthetics elements, and
 1. one or more geoms to visualize the data in the mappings

```{r}
penguins %>% 
  ggplot()
```

```{r}
penguins %>% 
  ggplot(aes(x = bill_length_mm, 
             y = bill_depth_mm))
```

```{r}
penguins %>% 
  ggplot(aes(x = bill_length_mm, 
             y = bill_depth_mm)) + 
  geom_point()
```

```{r}
penguins %>% 
  ggplot(aes(x = bill_length_mm, 
             y = bill_depth_mm)) + 
  geom_point() + 
  geom_smooth(method = "lm" )
```
What can be concluded with the visualization of the correlation between the bill depth and bill length?

### The Simpson Paradox
Add more layers of information

```{r}
penguins %>% 
  ggplot(aes(x = bill_length_mm, 
             y = bill_depth_mm)) + 
  geom_point() + 
  geom_smooth(method = "lm" ) + 
  aes(color = species)
```

## Data Wrangling

```{r}
penguins %>% 
  select(species)
```

```{r}
penguins %>% 
  filter(species == "Adelie")
```


## DISTRIBUTIONS

### Distribucions: geom_histogram
What is the weight of the penguins?
```{r}
penguins %>% 
  ggplot(aes(x = body_mass_g)) + 
  geom_histogram()
```

```{r}
penguins %>% 
  ggplot(aes(x = body_mass_g)) + 
  geom_histogram(bins = 10)
```

```{r}
penguins %>% 
  ggplot(aes(x = body_mass_g)) + 
  geom_histogram(bins = 10) + 
  aes(color = species)
```

```{r}
penguins %>% 
  ggplot(aes(x = body_mass_g)) + 
  geom_histogram(bins = 10) + 
  aes(fill = species)
```

### Facetting

```{r}
penguins %>% 
  ggplot(aes(x = body_mass_g )) + 
  geom_histogram(bins = 10) + 
  aes(fill = species) + 
  facet_wrap(~species)
```
More control

facet_grid

```{r}
penguins %>% 
  ggplot(aes(x = body_mass_g )) + 
  geom_histogram(bins = 10) + 
  aes(fill = species) + 
  facet_grid(species ~ .)
```

```{r}
penguins %>% 
  ggplot(aes(x = body_mass_g )) + 
  geom_histogram(bins = 10) + 
  aes(fill = species) + 
  facet_grid(species ~ sex)
```

```{r}
penguins %>% 
  drop_na() %>% 
  ggplot(aes(x = body_mass_g )) + 
  geom_histogram(bins = 10) + 
  aes(fill = species) + 
  facet_grid(species ~ sex)
```

```{r}
penguins %>% 
  drop_na() %>% 
  ggplot(aes(x = body_mass_g )) + 
  geom_histogram(bins = 10) + 
  aes(fill = species) + 
  facet_grid(sex ~ species)
```


###EXERCISE DATA EXPLORATION AND HISTOGRAMS 

Load NHANES
```{r}
NHANES
```
```{r}
glimpse(NHANES)
```
Explore NHANES

Make some tables for Gender

Table for Race1

Crosstable Gender and Race1

Create a new dataset for 2011_12
```{r}
nhanes_2011 <- NHANES %>% filter(NHANES$SurveyYr == "2011_12" )
```

Explore the distribution of BMI

```{r}
nhanes_2011 %>% 
  ggplot(aes(x = BMI)) + 
  geom_histogram()
```
Change the bins

```{r}
nhanes_2011 %>% 
  ggplot(aes(x = BMI)) + 
  geom_histogram(bins = 10)
```
Divide by gender

```{r}
nhanes_2011 %>% 
  ggplot(aes(x = BMI)) + 
  geom_histogram(bins = 10) + 
  facet_wrap(~Gender)
```
By sex and race

```{r}
nhanes_2011 %>% 
  ggplot(aes(x = BMI)) + 
  geom_histogram(bins = 30) + 
  facet_grid(Gender~ Race1)
```

### Distribucions: geom_bar

Which island do penguins come from?

```{r}
penguins %>% 
  ggplot(aes(x = island)) + 
  geom_bar() + 
  facet_wrap(~species)
```
```{r}

```



## ASSOCIATIONS

### Distribucions: geom_point

```{r}
penguins %>% 
  ggplot(aes(x = bill_length_mm, 
             y = bill_depth_mm)) + 
  geom_point() + 
  geom_smooth(method = "lm" ) + 
  aes(color = species)
```

Explore the correlation of height and weight for the NHANES dataset

```{r}
nhanes_2011 %>% 
  ggplot(aes(x = Height, 
             y = Weight)) + 
  geom_point()
```

```{r}
nhanes_2011 %>% 
  ggplot(aes(x = Height, 
             y = Weight)) + 
  geom_point() + 
  aes(color = Gender)
```

```{r}
nhanes_2011 %>% 
  ggplot(aes(x = Height, 
             y = Weight)) + 
  geom_point() + 
  aes(color = Gender) + 
  facet_wrap(~Gender)
```

```{r}
nhanes_2011 %>% 
  ggplot(aes(x = Height, 
             y = Weight)) + 
  geom_point() + 
  aes(color = Gender) + 
  facet_wrap(~Race1) 
```


BMI by hours sleep

```{r}
nhanes_2011 %>% 
  ggplot(aes(x = BMI, 
             y = SleepHrsNight)) + 
  geom_point() +
  geom_smooth()
```
Head circumference and HHIncomeMid

```{r}
nhanes_2011 %>% 
  ggplot(aes(x = HeadCirc, 
             y = HHIncomeMid, 
             color = Race1)) +
  geom_point()
```


and by gender?

```{r}
nhanes_2011 %>% 
  ggplot(aes(x = HeadCirc, 
             y = HHIncomeMid, 
             color = Race1)) +
  geom_point() + 
  geom_smooth(method = "lm")
```


## COMPARISONS

One variable is factor and the another is continuous


Example: boddy_mass by species
```{r}
penguins %>% 
  ggplot(aes(x = species, 
             y = body_mass_g)) + 
  geom_boxplot()
```
Explore the NHANES and identify differences in TotChol

```{r}
nhanes_2011 %>% 
  ggplot(aes(x = Race1, 
             y = TotChol)) + 
  geom_boxplot()
```
and now by gender: hint: add a new aes as color or fill

```{r}
nhanes_2011 %>% 
  ggplot(aes(x = Race1, 
             y = TotChol)) + 
  geom_boxplot() + 
  aes(color = Gender)
```


## TRENDS

```{r}
covid <- read_csv("https://opendata.ecdc.europa.eu/covid19/casedistribution/csv/data.csv")
```
```{r}
head(covid)
```


