---
title: "3rd Session: Data wrangling"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---


```{r}
pacman::p_load(tidyverse, gapminder)
```

# Dataset

```{r}
# Look at the gapminder dataset
gapminder
```

How many observations (rows) are in the dataset?

## Filter

Add a filter() line after the pipe (%>%) to extract only the observations from the year 1957. Remember that you use == to compare two values.

```{r}
# Filter the gapminder dataset for the year 1957
gapminder %>%
  filter(year == 1957)
```

```{r}
# Filter for China in 2002
gapminder %>%
  filter(year == 2002, country == "China")
```

## Arrange

```{r}
# Sort in ascending order of lifeExp
gapminder %>% arrange(lifeExp)
  

```
```{r}
# Sort in descending order of lifeExp
gapminder %>% arrange(desc(lifeExp))
```

Use filter() to extract observations from just the year 1957, then use arrange() to sort in descending order of population (pop).


```{r}
# Filter for the year 1957, then arrange in descending order of population
gapminder %>% filter(year == 1957) %>% arrange(desc(pop))
```

## Mutate

Suppose we want life expectancy to be measured in months instead of years: you'd have to multiply the existing value by 12. You can use the mutate() verb to change this column, or to create a new column that's calculated this way.

```{r}
# Use mutate to change lifeExp to be in months
gapminder %>% 
        mutate(lifeExp = lifeExp * 12)
```


```{r}
# Use mutate to create a new column called lifeExpMonths
gapminder %>% 
        mutate(lifeExpMonths = lifeExp * 12)
```

In this exercise, you'll combine all three of the verbs you've learned in this chapter, to find the countries with the highest life expectancy, in months, in the year 2007.

In one sequence of pipes on the gapminder dataset:
filter() for observations from the year 2007,
mutate() to create a column lifeExpMonths, calculated as 12 * lifeExp, and
arrange() in descending order of that new column

```{r}
# Filter, mutate, and arrange the gapminder dataset
gapminder %>%
  filter(year == 2007) %>%
  mutate(lifeExpMonths = 12 * lifeExp) %>%
  arrange(desc(lifeExpMonths))
```


# Visualizing with ggplot2

```{r}
# Create gapminder_1952
gapminder_1952 <- gapminder %>% filter(year == 1952)
gapminder_1952
```

Change the scatter plot of gapminder_1952 so that (pop) is on the x-axis and GDP per capita (gdpPercap) is on the y-axis.

```{r}
# Change to put pop on the x-axis and gdpPercap on the y-axis
gapminder_1952 %>% 
  ggplot(aes(x = pop, y = gdpPercap)) +
  geom_point()
```
Create a scatter plot of gapminder_1952 with population (pop) is on the x-axis and life expectancy (lifeExp) on the y-axis.
```{r}
# Create a scatter plot with pop on the x-axis and lifeExp on the y-axis
ggplot(gapminder_1952, aes(x = pop, y = lifeExp)) +
        geom_point()
```

## Log scales
```{r}
# Change this plot to put the x-axis on a log scale
ggplot(gapminder_1952, aes(x = pop, y = lifeExp)) +
  geom_point() +
  scale_x_log10()
```
Create a scatter plot with population (pop) on the x-axis and GDP per capita (gdpPercap) on the y-axis. Put both the x- and y- axes on a log scale.

```{r}
# Scatter plot comparing pop and gdpPerCap, with both axes on a log scale
gapminder_1952 %>% 
        ggplot(aes(x = pop, y = gdpPercap)) + 
        geom_point() + 
        scale_x_log10() + scale_y_log10()
```
Create a scatter plot with population (pop) on the x-axis, life expectancy (lifeExp) on the y-axis, and with continent (continent) represented by the color of the points. Put the x-axis on a log scale.

```{r}
# Scatter plot comparing pop and lifeExp, with color representing continent
gapminder_1952 %>% 
        ggplot(aes(x = pop, y = lifeExp, color = continent)) +
        geom_point() + 
        scale_x_log10()
```
Modify the scatter plot so that the size of the points represents each country's GDP per capita (gdpPercap).

```{r}
# Add the size aesthetic to represent a country's gdpPercap
ggplot(gapminder_1952, aes(x = pop, y = lifeExp, color = continent, size = gdpPercap)) +
  geom_point() +
  scale_x_log10()
```
## Facetting

Create a scatter plot of gapminder_1952 with the x-axis representing population (pop), the y-axis representing life expectancy (lifeExp), and faceted to have one subplot per continent (continent). Put the x-axis on a log scale.

```{r}
# Scatter plot comparing pop and lifeExp, faceted by continent
gapminder_1952 %>% 
  ggplot(aes(x = pop, y = lifeExp)) + 
  geom_point() + 
  scale_x_log10() + 
  facet_wrap(~continent)
```
Create a scatter plot of the gapminder data:
Put GDP per capita (gdpPercap) on the x-axis and life expectancy (lifeExp) on the y-axis, with continent (continent) represented by color and population (pop) represented by size.
Put the x-axis on a log scale
Facet by the year variable

```{r}
# Scatter plot comparing gdpPercap and lifeExp, with color representing continent
# and size representing population, faceted by year
gapminder %>% 
        ggplot(aes(x = gdpPercap, y = lifeExp, 
                   color = continent, size = pop)) +
        geom_point() + 
        scale_x_log10() + 
        facet_wrap(~year)
```
# Summarizing

Use the median() function within a summarize() to find the median life expectancy. Save it into a column called medianLifeExp.
